# Thurs Sept 24: Protection Island 3 {#PE3}

DUE DATE: Thurs Oct 1

Here is some additional information also taken from:
Newcomb, HR. 1940. [Ring-necked pheasant studies on Protection Island in the Strait of Juan de Fuca](https://ir.library.oregonstate.edu/concern/graduate_thesis_or_dissertations/js956j801?locale=en), Washington. MS thesis. Oregon State University.

a.	Pheasant chicks are born during the summer.
b.	In May 1937, 10 pheasants were introduced to the island. Before the next breeding season there were 35.
c.	November 10, 1938 a census estimated 110 pheasants.
d.	October 13, 1939 a census estimated 400 pheasants.
e.	Between the 1938 and 1939 censuses, Newcomb observed that 17 adult birds died.
f.	During the 1938 nesting season there were 5.86 eggs/nest. 83.57% of eggs hatched.
g.	During the 1939 nesting season there were 8.73 eggs/nest. 64.58% hatched.
h.	During the 1939 nesting season: Average number of chicks per clutch was 6.93.$^1$ 
i.	You can assume the sex ratio is 50:50 male to female. Pheasants are a sexually reproducing species.

$^1$ Note that g. and h. appear to be contradictory.

## Questions

1.	Let $d$ be the fraction of population that dies each year. Estimate $d$ for the ring-tailed pheasant population on Protection Island? Write down any assumptions you have made. [3 marks]

2.	$b$ is the per capita number of births each year. What is the value of $b$? Write down any assumptions you have made. [3 marks]

3.	Recall that $\lambda = 1 + b-d$. What is the value of $\lambda$? Is this population is expected to grow over time? [2 marks]

4.	Lets assume that the pheasant population on Protection Island grows geoemetrically (i.e. exponentially but for a discrete time model) where the geometric growth rate, $\lambda$, is the value that you estimated in 3. Lets predict the population size each May beginning with May 1937. Let $N_0 = 10$ and let $t$ be the number of years since May 1937. Recall that when a population grows geometrically,

\[N_{t} = N_0 \lambda^t \]

You can use `R` to do this calculation as follows (you should use your value of $\lambda$ from question 5, rather than `lambda <- 3` as in the example below):

```{r,eval=FALSE}
t <-1
N0 <-10
lambda <-3
N0*lambda^t
```

The result of `N0*lambda^t` is $N_{t+1}$, and with $t=1$ then $N_{t+1}=N_2$: the population size two years after May 1937. You can change the value of $t$ and repeat the calculation. Unless you have cleared your workspace it won't be necessary to re-input $N_0 = 10$ and $\lambda = 3$. As such, you can calculate $N_3$ with the following commands:

```{r,eval=FALSE}
t <-2
N0*lambda^t
```

<strong> HAND IN </strong>

Use `R` to predict the value of the pheasant population size every year up until May 1940. You only need to hand in the values that you get, not an R Script. [2 marks]

The approach to calculating the pheasant population size in Question 4 is not very organized. In this question, we will learn how to make a data frame, use a for loop, and use the function `rbind()`.


Read [Data structures]( https://ahurford.github.io/quantitative-training-guide/rintro.html#data-structures) in _Quantitative training for biology_.

Create a one row dataframe called `df`:

```{r,eval=FALSE}
df <- data.frame(time = 0, popn.size = 10)
```

Query `df` in your `Console` to see the data frame you have created. We would like to add successive values of the population size that we calculate to the data frame. To do this we use the `rbind()` function, which binds rows together.

```{r,eval=FALSE}
new.result <- data.frame(time = 1, popn.size = 20)
df <- rbind(df, new.result)
```

Here the `rbind()` function takes the `df` dataframe and adds the `new.result` data frame as a new row onto the bottom. Note that the code above _overwrites_ the value of `df`: that is, `new.result` is added to the bottom of the `df` dataframe (containing only one row), and the result is called `df` (which now has two rows), and the old dataframe `df` (with one row) is overwritten. As such, each time you run the command `df <- rbind(df, new.result)` another row is added to `df`. Try the following:

```{r,eval=FALSE}
new.result <- data.frame(time = 1, popn.size = 20)
df <- rbind(df, new.result)
df <- rbind(df, new.result)
df <- rbind(df, new.result)
```

If you query the value of `df` you can see that the several rows, all with identical values have been added because we have run the command `df <- rbind(df, new.result)` multiple times while the value of `new.result` is unchanged. Now let's change the value of `new.result` between each time we run the `df <- rbind(df, new.result)` command.

```{r,eval=FALSE}
new.result <- data.frame(time = 1, popn.size = 20)
df <- rbind(df, new.result)
new.result <- data.frame(time = 2, popn.size = 30)
df <- rbind(df, new.result)
```

Type `df` into the `Console` to see the resulting dataframe. Finally, when we do calculations for a sequence of values, it is easier to code this using a `for` loop.

```{r,eval=FALSE}
lambda <- 1.2
N0 <- 10
df <- data.frame(time = 0, popn.size = 10)
for(t in seq(1,4,1)){
  val <- N0*lambda^t
  new.result <- data.frame(time = t, popn.size = val)
  df <- rbind(df, new.result)
}
```

To understand the above code, after copy and pasting it into your `Console`, and running it by clicking `Return`, query the value of `df`: you should see predicted population sizes up until 4 years after May 1937. Now, lets try to understand `seq(1,4,1)`. Let's learn about the `seq()` function by trying it out in the `Console`. What is the result of each of these?

```{r,eval=FALSE}
seq(-10,10)
seq(-10,5,0.1)
```

The `for` loop works by beginning with `t` equal to the first value of the sequence and stepping through each value until the final value. The code is written so that quantities that depend on `t` are inside the `for` loop (i.e., enclosed with in the `{}` and those that do not depend on `t` are outside the `for` loop). Note that `val` changes for different values of `t`, `new.result` changes for different values of `t` (because `new.result` has `time = t` and `pop.size = val`, where `val` depends on `t`). Finally, `df` also depends on `t`, because `new.result` depends on `t`. In contrast, `N0` and `lambda` do not change with `t`, so it is more efficient to place the allocated values for these parameters outside of the loop.

We can also plot the results of our calculations:

```{r}
lambda <- 2
N0 <- 10
df <- data.frame(time = 0, popn.size = 10)
for(t in seq(1,4,1)){
  val <- N0*lambda^t
  new.result <- data.frame(time = t, popn.size = val)
  df <- rbind(df, new.result)
}
plot(df$time, df$popn.size, typ = "l", xlab = "years since May 1937", ylab = "Population size")
```

If you already have an existing plot you can add new lines using `lines()`. For example,


```{r}
plot(seq(1,4), c(1,3,4,2), ylab = "y-axis", xlab = "x-axis")
lines(seq(1,4), seq(1,4))
```

<strong> HAND IN </strong>

5. Write an R scipt that builds on the file you have previously made _protection-island.R_. Use the `lines()` command to add the predicted population size assuming geometric growth using the commands described in this section. If you have written the code correctly the result should look something like this:

```{r,eval=T,echo=F,message=F}
# Title: Protection Island – chron package
# Author: Amy Hurford
# Date: Sept 16, 2020

# Load package
require(chron)

# Population sizes
y <- c(10,35,110,440)

# Dates
ref.day <-c(5,1,1937)
x <- c(julian(5,1,1937, origin = ref.day), julian(5,1,1938, origin = ref.day), julian(11,10,1938, origin = ref.day), julian(10,13,1939, origin = ref.day))

# Plot
plot(x/365,y, xlab = "years since May 1, 1937", ylab = "no. of pheasants")

lambda <- 3.48
N0 <- 10
df <- data.frame(time = 0, popn.size = 10)
for(t in seq(1,4,1)){
  val <- N0*lambda^t
  new.result <- data.frame(time = t, popn.size = val)
  df <- rbind(df, new.result)
}
lines(df$time, df$popn.size, typ = "l", xlab = "days since May 1937", ylab = "Population size")
```

Note that to get years on the x-axis for data from the _protection-island.R_ file, I needed to plot `plot(x/365,y)` because with `plot(x,y)` the units on my x-axis were days. Your graph may look slightly different since you may have calculated a slightly different $\lambda$ value if you made different assumptions to me in questions 1 and 2.

You are to hand in your R Script.  See \@ref(RScript) for more information. [10 marks]

